\documentclass{article}

\usepackage{fullpage}
\usepackage{setspace}
\usepackage{mathtools}

\title{Detecting Deadlock in Queueing Network Simulations}
\author{Geraint Palmer}
\date{}

\begin{document}
\onehalfspacing

\maketitle

\section{Explanation of Deadlock}
Any open queueing network with feedback loops, at least one service station that has limited queueing capacity, and where individuals can be blocked from joining the queue at the next destination can experience deadlock. Deadlock occurs when a customer finishes service at node $i$ and is blocked from transitioning to node $j$; however the individuals in node $j$ are all blocked, directly or indirectly, by the blocked individual in node $i$.
NEED A LOT MORE HERE. DIAGRAMS TO EXPLAIN TOO.

\section{Literature Review}
Most of the literature on blocking conveniently assumes the networks are deadlock-free. For closed networks of $K$ customers with only one class of customer, \cite{kunduakyildiz89} proves the following condition to ensures no deadlock: for each minimum cycle $C$, $K < \sum_{j\in C} B_j$, the total number of customers cannot exceed the total queueing capacity of each minimum subcycle of the network. The paper also presents algorithms for finding the minimum queueing space required to ensure deadlock never occurs, for closed cactus networks, where no two cycles have more than one node in common. This result is extended to multiple classes of customer in \cite{liebeherrakyildiz95}, with more restrictions such as single servers and each class having the same service time distribution. Here a integer linear program is formulated to find the minimum queueing space assignment that prevents deadlock. The literature does not discuss deadlock properties in open restricted queueing networks.\newline
NEED A LOT MORE HERE

\section{Definitions}
\begin{tabular}{ p{5cm} p{10cm} }
  $\left| V(D) \right|$ & The order of the directed graph $D$ is its number of vertices. \\
  Weakly connected component & A weakly connected component of a digraph containing $X$ is the set of all nodes that can be reached from $X$ if we ignore the direction of the edges. \\
  Direct successor & If a directed graph contains an edge from $X_i$ to $X_j$ , then we say that $X_j$ is a direct successor of $X_i$. \\
  Ancestors & If a directed graph contains a path from $X_i$ to $X_j$ , then we say that $X_i$ is an ancestor of $X_j$. \\
  Decendents & If a directed graph contains a path from $X_i$ to $X_j$ , then we say that $X_j$ is a descendant of $X_i$. \\
  $deg^{out}(X)$ & The out-degree of $X$ is the number of outgoing edges emanating from that vertex. \\
\end{tabular}
\newline
NEED CONSISTANT NOTATION, REFERENCES FOR DEFINITIONS.

\section{Explaining Digraph}
Here is a method of detecting when deadlock occurs in an open queueing network $Q$ with $N$ nodes. Let the number of servers in node $i$ be denoted by $c_i$.\newline
Define $D$ as a directed graph associated with $Q$, where $\left| V(D) \right| \leq \sum_{i=1}^N c_i$. Each vertex of $D$ is an individual who is either in service or blocked.\newline
When an individual finishes service at node $i$, and this individuals next destination is node $j$, but there is not enough queueing capacity for $j$ to accept that individual, then that individual remains at node $i$ and becomes blocked. At this point $c_j$ directed edges between this individual and each individual in service or blocked at node $j$ are created in $D$.\newline
Once an individual is released, that individual is removed as a vertex of $D$, and all edges in and out of that vertex removed. If another individual enters service at the same time, then that next individual node, and aquires the previous individual's in-edges.\newline
CLEARER, VERTICES SHOULD BE SERVERS THAT CAN BE OCCUPIED. LOTS OF DIAGRAMS. SUPGRAPHS THING.

\section{Observations}
Consider one weakly connected component of $D$, $G$. Consider the node $X_a \in G$, where $X_a$'s next destination is node $j$. Then $X_a$'s direct successors are the individuals who are blocked or in service at node $j$. We can interpret all $X_a$'s decendents as those individuals who are directly or indirectly blocking $X_a$, and we can interpret all $X_a$'s ancestors as those individuals who are being blocked directly or indirectly by $X_a$.\newline
We do not need to worry about the situation $deg^{out}(X_a) > c_j$ as it will never occur because we only create edges from $X_a$ to individuals in service or blocked in node $j$. If $deg^{out}(X_a) = c_j$ then $X_a$ is blocked by all its direct successors. The only other situation is that $X_a$ is not blocked, and belongs to $G$ because $X_a$ is in service and blocking other individuals, in which case $deg^{out}(X_a) = 0$.\newline
It is clear that if any of $X_a$'s decendents are not blocked, then we do not have deadlock, and if all of $X_a$'s descendents are blocked, then we have deadlock. We also know that by definition all of $X_a$'s ancestors are blocked. It is therefore also clear that, for any individual $X$, if all of $X$'s descendants are blocked, then the descendents of every node in $G$ are blocked. Therefore to check that we haven't got deadlock, we only need to check that there is at least one node of $G$ that is not blocked; hence checking for a node in $G$ that has $deg^{out} = 0$.\newline

\section{Theorem}

\subsection*{Theorem}
A deadlock situation arises if and only if there exists a weakly connected component of $D$ containing no vertices with $deg^{out} = 0$.\newline

\subsection*{Observations}
Consider one weakly connected component $G$ of $D$. Consider the node $X_a \in G$, where $X_a$'s next destination is node $j$. Then $X_a$'s direct successors are the individuals who are blocked or in service at node $j$. We can interpret all $X_a$'s decendents as those individuals who are directly or indirectly blocking $X_a$, and we can interpret all $X_a$'s ancestors as those individuals who are being blocked directly or indirectly by $X_a$.\newline
We do not need to worry about the situation $deg^{out}(X_a) > c_j$ as it will never occur because we only create edges from $X_a$ to individuals in service or blocked in node $j$. If $deg^{out}(X_a) = c_j$ then $X_a$ is blocked by all its direct successors. The only other situation is that $X_a$ is not blocked, and belongs to $G$ because $X_a$ is in service and blocking other individuals, in which case $deg^{out}(X_a) = 0$.\newline
It is clear that if any of $X_a$'s decendents are not blocked, then we do not have deadlock, and if all of $X_a$'s descendents are blocked, then we have deadlock. We also know that by definition all of $X_a$'s ancestors are blocked.\newline

\subsection*{Proof}
Consider one weakly connected component $G$ of $D$.\newline
Assume that $G$ contains a vertex $X$ such that $deg^{out}(X) = 0$. This imples that $X$ is not blocked, therefore $Q$ is not deadlocked.\newline
Now assume that we have deadlock, and all of the members of $G$ are blocked. For a given vertex $X$, all decendents of $X$ are blocked, and so must have out-degrees greater than 0. However, as our choice of $X$ was arbitrary, this must be true for all members of $G$, and no members of $G$ have out-degree of 0.\newline

\bibliographystyle{plain}
\bibliography{refs}





\end{document}
